---
title: 'Task 1: Dissolving NYS County Shape Files'
author: "MW"
output: 
  html_document:
    toc: true
    theme: cosmo
---


```
Task: I want to dissolve the five counties in NYC: Queens, Kings, Bronx, New York, Richmond to be one polygon called "NYC" 
```

- __Data__: The New York State County Shape File was downloaded [here]( https://cugir.library.cornell.edu/catalog/cugir-007865)
- __R libraries__:
  - [sf: geo process tools in R](https://github.com/r-spatial/sf)
  - [tidyverse: data manipulation tools (dplyr)](https://www.tidyverse.org)
  - [leaflet: plot maps in R](https://rstudio.github.io/leaflet/)


### Step 1: Loading Packages and Data 
- plot the shape file to check if the shape file is properly loaded
- for shape file to load properly, it is important to keep all files in the folder available (.dbf, .prj, .sbn, .shp etc). 

```{r}
require(sf)
require(tidyverse)
require(leaflet)

# nys shape file 
nys_shape <- st_read(
  dsn = "data/cugir-007865/",
  layer = "cty036"
)

leaflet(nys_shape) %>% addPolygons() # checking
```


### Step 2. Dissolve NYC Region Counties 
- key steps: only extract the NYC Regions and use function `st_union` to dissolve the entire area 
```{r}
# get the nyc part only 
nyc_shape <- nys_shape %>% filter(NAME %in% c("Bronx", "Queens", "Kings", "New York", "Richmond")) %>% 
  distinct(NAME, COUNTY, geometry)

leaflet(nyc_shape) %>% addPolygons() # checking 


# dissolve nyc 
nyc_shape_dissolved <- st_union(nyc_shape) %>% 
  st_sf() %>% 
  cbind(NAME = "New York City", COUNTY = "061")

leaflet(nyc_shape_dissolved) %>% addPolygons() # checking
```

### Step 3. Paste the NYC Region with the "rest of state" counties

```{r}
# get the rest of state part 
ros_shape <- nys_shape %>% 
  filter(!NAME %in% c("Bronx", "Queens", "Kings", "New York", "Richmond")) %>% 
  distinct(NAME, COUNTY, geometry)
leaflet(ros_shape) %>% addPolygons() # checking 


# union with ros 
nys_dissolved <- rbind(st_sf(ros_shape), 
                       st_sf(nyc_shape_dissolved))

leaflet(nys_dissolved) %>% addPolygons() # checking
```

